Automated Topic Naming

### 2011-08-25 ese/ese.tex
neil's changes to ese.tex

To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level NFRs: maintainability, functionality,
portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
seems like a reasonable starting point for our work.}.
%There is some debate about the terms in this model. 
%However, ISO9126 is ``an international standard and thus provides an 
%internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
projects.

%\noindent \textbf{Creating a validation corpus} ---

\subsubsection{Creating a Validation Corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For MySQL 3.23
and MaxDB 7.500, the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label. Sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. 
%It is really important and it is required by both analyses. Plus we can tell them how.}

@@

%\subsection{Generating word lists}
%\label{sec:genwordlist}
% %\todo[inline]{Make it clear where the data comes from! Neil says: Also, I
%   didn't get the impression they understood that we weren't generating
%   labels from the data, but rather from independent taxonomies. Of
%   course if you use labels from the data it isn't comparable. But the
%   point of ISO 9126 etc. was to avoid exactly this problem.}

% \todo[inline]{Need to hammer it in that the word bags are not project specific and
%  are inherited so it isn't a supervised method}

%Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
%While the topics themselves are generated automatically, how to interpret these topics is less clear.
%For example,

@@

%\noindent\textbf{Generating word lists} ---

@@

%\todo[inline]{* clarify the mailinglist stuff we did}
%\todo[inline]{This part doesn't make a lot of sense}
To construct these expanded word-lists, we used
WordNet~\cite{Fellbaum1998}.
%, an English-language ``lexical database'' that contains semantic relations between words,
% including common related forms (similar to word stemming), meronymy and synonymy. 
We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

@@

%\todo[inline]{NEIL: this could be a subsubsection}
%\todo[inline]{* the hierarchy of tags was not clear}
%\noindent \textbf{Generating labels automatically} ---

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a matching NFR label. \emph{Unnamed
topics} occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional
requirements.
All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
extracted $20$ topics per period for each project.
% Is this good enough
This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
the project being analyzed, and we did not label the project's topics
ourselves for a training set. The motivation behind this technique is that
because most software often addresses similar issues, thus we can use the
domain knowledge of software to label relevant topics.
%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}

>>>>

To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level NFRs: maintainability, functionality,
portability, efficiency, usability, and reliability~\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
seems like a reasonable starting point for our work.}.
There is some debate about the terms in this model. 
However, ISO9126 is ``an international standard and thus provides an 
internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
projects.

\subsubsection{Creating a Validation Corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For each project, 
the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label. Sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. 
%It is really important and it is required by both analyses. Plus we can tell them how.}

@@

@@

@@

To construct these expanded word-lists, we used
WordNet~\cite{Fellbaum1998}.
%, an English-language ``lexical database'' that contains semantic relations between words,
% including common related forms (similar to word stemming), meronymy and synonymy. 
We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

@@

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a matching NFR label. \emph{Unnamed
topics} occur where there is no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional
requirements.
All experiments were run on the datasets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
extracted $20$ topics per period for each project.
This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
the project being analyzed, and we did not label the project's topics
ourselves for a training set. The motivation behind this technique is that
because most software often addresses similar issues, we can use the
domain knowledge of software to label relevant topics.
